\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{fullpage}
\usepackage{mpemath}
\usepackage{hyperref} % For hyperlinks in the PDF like citations
\usepackage{dsfont}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[capitalise,noabbrev]{cleveref}
\usepackage{algorithm,algorithmicx,algpseudocode}
\usepackage{bm}

\newcommand{\gersi}[1]{\textcolor{red}{#1}}
\newcommand{\palash}[1]{\textcolor{green}{#1}}
\newcommand{\mm}[1]{\textcolor{magenta}{[#1]}}
\newcommand{\maeve}[1]{\textcolor{blue}{[#1]}}
\bibliographystyle{abbrvnat}

\DeclareMathOperator{\kl}{KL}
\DeclareMathOperator{\ess}{ess}
\DeclareMathOperator{\KL}{KL}
\DeclareMathOperator{\ext}{ext}
\DeclareMathOperator{\conv}{conv}
\newcommand{\quant}{{\mathfrak{q}}}

%\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
%\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\DeclarePairedDelimiter{\spannorm}{\lVert}{\rVert_s}
% \newtheoremstyle{case}{}{}{}{}{}{:}{ }{}
% \theoremstyle{case}
% \newtheorem{case}{\textbf{Case}}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,automata}

\DeclareMathOperator{\varo}{VaR}
\DeclareMathOperator{\cvaro}{CVaR}
\DeclareMathOperator{\evaro}{EVaR}

%\title{CVaR in Linear Time}
\title{Fast Algorithms for VaR and $\varphi$-divergence Risk Measures, like CVaR and EVaR}
\author{Palash Agrawal, Gersi Doko, Maeve Burwell, Marek Petrik}
\date{}

\begin{document}

\maketitle

\section{Introduction}
% https://www.researchgate.net/publication/2611689_A_One-Pass_Space-Efficient_Algorithm_for_Finding_Quantiles
Approximate quantile algorithm but probability distribution is not known unlike our setting (really each element in the given dataset has an equal probability)


% https://arxiv.org/pdf/2310.07224
(Top-k projection - note this is DIFFERENT than k-capped simplex because it does not have constraint that elements should be between 0 and 1 and it may not necessarily be looking at the sum of all elements, just the top k elements). The paper claims that this can be trivially extended to the superquantile (CVaR), however this mainly seems to be the case other then if you are sampling from an unknown continuous distribuiton. That being said, they have a Rockefellar-Ursalyev formulation (well known in Cvar literature) formulation to the top-k operator in their prior work. They also have a quickselect formulation to solve this problem

% https://arxiv.org/pdf/1503.01002
A sorting and partitioning based approach to the k-capped simplex problem (with the constraint being the sum of the vector has to add up to k). Simple proof done with the KKT conditions.

% https://hal.science/hal-01056171v2/document
Excellent paper covering the k-capped simplex problem in two forms (when the sum can be less than dor equal to k or when it can be equal to k). Covers lots of prior algorithms used to solve the problem (I need to read this in more depth). This paper has a lemma that will come helpful in equating the inequality constraint in k-capped simplex problem to the equality of constraint of the sum of probability vector equal to 1 in Cvar case.

% https://bayen.berkeley.edu/sites/default/files/efficient_bregman_projections.pdf
This paper seems to describe an algorithm for projection onto the unit simplex, NOT k-capped using a quickselect based approach, although this is not the primary contribution. 


\section{Preliminaries}

Before describing CVaR and its solution we define the basic notation we use in the paper. 
We use calligraphic letters to denote sets and a tilde to denote random variables. 
We also adopt the standard convention that $\mathcal{A}^{\mathcal{B}}$ represents the set of all functions from a set $\mathcal{B}$ to a set $\mathcal{A}$ and 
treat vectors as a function from indexes to real numbers. $\mathbf{1}$ denotes the vector of ones and $\mathds{1}(bool)$ denotes the indicator function. We use the convention that comparison operators are defined on vector spaces element-wise, that is for $x, y \in \mathbb{R}^n$ then $x \leq y \iff x_i \leq y_i \quad \; \forall \; i \in 1\dots n$.
Finally, $\Delta^n$ denotes the standard $n$-simplex while $\Delta_k^n$ denotes the k-capped simplex~\cite{kcapped_ang}.

\section{Formal Model}


\paragraph{Value-at-Risk and quantiles}
Quantile of a random variable $\tilde{x} \in \mathbb{X}$ at level $\alpha \in [0,1]$ is any value $\tau\in \mathbb{R}$ such that $\mathbb{P}({\tilde{x}) \le \tau} \ge \alpha$ and $\mathbb{P}({\tilde{x}) \ge \tau} \ge 1-\alpha$. The quantile of a random variable may not be unique and lies in the interval $[\quant^-_{\alpha}(\tilde{x}), \quant^+_{\alpha}(\tilde{x})]$ where
\begin{align} \label{eq:quantile-definition}
%\label{eq:upper-quantile-definition}
  \quant_{\alpha}^+(\tilde{x})
  &\;:=\;
    \max\; \left\{\tau \in \bar{\mathbb{R}} \mid \mathbb{P}({ \tilde{x}) < \tau } \le \alpha\right\}, 
      %\label{eq:low-quantile-definition}
  &\quant_{\alpha}^-(\tilde{x})
  &\;:=\;
    \min\; \left\{\tau \in \bar{\mathbb{R}} \mid \mathbb{P}({ \tilde{x})  \le \tau } \ge \alpha\right\}.
\end{align}
The maximum in \eqref{eq:quantile-definition} exists because $\tau \mapsto \mathbb{P}({\tilde{x} )< \tau}$ is lower semi-continuous and the minimum in \eqref{eq:quantile-definition} exists because $\tau \mapsto \mathbb{P}({\tilde{x}) \le \tau}$ is upper semi-continuous. Note that $\quant^+_1(\tilde{x}) = \infty$, $\quant^-_0(\tilde{x}) = -\infty$ and the quantiles are bounded otherwise for all $\alpha \in [0,1]$.

Monetary risk measures generalize the expectation operator to account for the uncertainty in the distribution. \emph{Value-at-Risk}~(VaR) is one of the oldest, simplest, and most common risk measures~\cite{Follmer2016, Shapiro2014}. Value-at-Risk: $\varo \colon \mathbb{X} \to \bar{\mathbb{R}}$ at a risk level $\alpha \in [0,1]$ of a random variable $\tilde{x}\in \mathbb{X}$ is defined as the largest $1-\alpha$ confidence lower bound on the value of the random variable:
\begin{equation}
  \label{eq:var-definition}
\varo_{\alpha}{\tilde{x}} \;:=\;  \quant^+_{ \alpha }(\tilde{x}).  
\end{equation}
\paragraph{$\varphi$-divergence Risk Measures}

\mm{Palash?}

\begin{enumerate}
\item Define $\varphi$-divergence
\item Define $\varphi$-divergence risk measures 
\cite{Ahmadi-Javid2017,Ahmadi-Javid2012}
\item Some more recent papers that use these risk measures \cite{Kruse2021}
\item CVaR, EVaR, VVaR as $\varphi$-divergence risk measures have a table something like Table 1 in \cite{Bayraksan2015}
\end{enumerate}

\paragraph{Polymatroids}
 This is something that we do want to mention, but maybe not.
\begin{enumerate}
\item Define the polymatroids
\item Define the generic algorithm
\end{enumerate}



We can assume a probability space with outcomes $\Omega = \left\{ 1, \dots , N \right\}$, events $\mathcal{F}$, and
probability function $\mathbb{P} \in \Delta^N$. We consider a random variable to be any mapping $\tilde{x}\colon \Omega \to \mathbb{R}$. Or equivalently as vectors on $\Omega$. For example, the vector $\bm{x}\in \mathbb{R}^N$ equivalent to $\tilde{x}$ is constructed as
\[
 x_w = \tilde{x}( \omega ), \quad\forall \omega \in \Omega .
\]
We assume the same equivalence for probability distributions. 

\section{Fast Algorithm for VaR}



\section{Fast $\varphi$-divergence Algorithms}

\subsection{General Algorithm}

\subsection{CVaR}
As observed in~\cite{rockefeller1996variational}, the Conditional Value-at-Risk (CVaR) is a convex risk measure that generalizes the Value-at-Risk (VaR) and can be definied as
\begin{equation} \label{eq:primal_cvar}
  \begin{aligned}
    \cvaro_\alpha(\tilde{x})
  &= \min \left\{  \Ex{\bm{\xi}}{ \tilde{x}  } \mid
  \bm{\xi} \leq \frac{1}{1-\alpha} p, \bm{\xi} \in \Delta^n \right\} \\
  &= \min \left\{\bm{\xi}^T \bm{x} \mid \bm{\xi} \leq \frac{1}{1-\alpha} \bm{p},
     \xi^T \mathbf{1} = 1, \bm{\xi} \geq 0\right\}\\
  &= \max \left\{ z - \frac{1}{1-\alpha} \Ex{\bm{p}}{z-x}_+ \mid z \in \mathbb{R}\right\}.
  \end{aligned}
\end{equation}

\subsection{EVaR}

\subsection{VVaR}


\section{To Be Sorted}

\begin{thm}
\Cref{alg:quick_cvar} is correct and runs in linear time.
\end{thm}

\begin{lem}
Suppose that $\tilde{x}\colon \Omega \to \mathbb{R}$ is a random variable with finite $\Omega$, and $\alpha\in [0,1)$. Then, there exist $\omega_1, \omega_2\in \Omega$ such that $\quant_{\alpha}^+(\tilde{x}) = \tilde{x}(\omega_1)$ and $\quant_{\alpha}^-(\tilde{x}) = \tilde{x}(\omega_2)$ for each $\alpha\in (0,1]$.
\end{lem}
\begin{proof}
Recall the definition:
 \[
  \quant_{\alpha}^+(\tilde{x})
  \;:=\;
    \max\; \left\{\tau \in \bar{\Real} \mid \P{ \tilde{x} < \tau } \le \alpha\right\}, 
 \] 
Assume the random variable $ x $ has $ n $ unique values, such that $ x_1 < x_2 < \cdots < x_n $.  We can assume this without loss of generality because VaR is law invariant~\cite{Follmer2016}.  

Suppose for the sake of contradiction, that there is no $\omega\in \Omega $ such that $\quant^+_{\alpha}(\tilde{x}) = \tilde{x}(\omega)$. That may mean:
\begin{enumerate}
\item[Case 1] $\exists x_{i}, x_{i + 1} \in \tilde{x}(\Omega)$ such that $\quant^+_{\alpha}(\tilde{x}) \in (x_{i}, x_{i + 1})$.
Let $\tau\opt $ is optimal in \eqref{eq:var-definition} (max definition) and therefore $\tau\opt  = \quant^+_{\alpha}(\tilde{x})  $. Then, let $\tau' = \tau\opt  + \frac{1}{2}(x_{i+1} - \tau\opt)$. Because $\tau' > \tau\opt $ and $\tau'$ is feasible in \eqref{eq:var-definition}, $\P{ \tilde{x} < \tau' } =  \P{ \tilde{x} < \tau\opt  }\le \alpha$, $\tau\opt $ cannot be optimal.
\item[Case 2] $\quant^+_{\alpha}(\tilde{x}) < \min \tilde{x}$ 
  
  Let $\tau\opt $ is optimal in \eqref{eq:var-definition} (max definition) and therefore $\tau\opt  = \quant^+_{\alpha}(\tilde{x})  $. Then, let $\tau' = \tau\opt + \frac{1}{2}(x_{1} - \tau\opt)$. Because $\tau' > \tau\opt $ and $\tau'$ is feasible in \eqref{eq:var-definition}, $\P{ \tilde{x} < \tau' } =  \P{ \tilde{x} < \tau\opt  }\le \alpha$, $\tau\opt $ cannot be optimal.

\item[Case 3]  $\quant^+_{\alpha}(\tilde{x}) > \max \tilde{x}$ 

  

  Let $\tau\opt $ is optimal in \eqref{eq:var-definition} (max definition) and therefore $\tau\opt  = \quant^+_{\alpha}(\tilde{x})$. Recall $\alpha \in [0,1)$, $\P{\tilde{x} < \tau\opt} = 1 \leq \alpha < 1$.
  
  % New Proof 2: Recall that $\alpha$ is less than 1 and  $\quant^+_{\alpha}(\tilde{x})$ is bounded for all admissible $\alpha$ values. Let $\tau\opt $ is optimal in \eqref{eq:var-definition} (max definition) and therefore $\tau\opt  = \quant^+_{\alpha}(\tilde{x})  $. Then, let $\tau' = \tau\opt + 1$. Because $\tau' > \tau\opt $ and $\tau'$ is feasible in \eqref{eq:var-definition}, $\P{ \tilde{x} < \tau' } =  \P{ \tilde{x} < \tau\opt  }\le \alpha$, $\tau\opt $ cannot be optimal.

  % Old Proof: Then $ P[x < \quant_{\alpha}^+(\tilde{x})] $ is greater than or equal to $ P[x < x_{n}] $. Remember that $\alpha$ must be less than 1 and $\quant_{\alpha}^+(\tilde{x})$ is bounded for all $\alpha$ less than 1 and greater than or equal to 0, therefore $ P[x < \quant_{\alpha}^+(\tilde{x})] $ is less than 1. Then, there must exist $ y $ in the range $ [\quant_{\alpha}^+(\tilde{x}), \infty) $ such that $ P[x = y] > 0 $ . This is impossible by definition. 
\end{enumerate}
\end{proof}

\begin{thm}
\Cref{alg:quick_quantile} computes $\quant^+$.
\end{thm}
\begin{proof}
 Loop Invariant: Before and after every iteration $\quant^+$ will remain in the array $x$ at an index $i$ such that $ front \leq i \leq back $.

At the loop beginning: Every element in $x$ has an index $i$ such $ front \leq i \leq back $, since $front=1$ and $back=n$. By Lemma 1, there exists an element with index $i$ such that $ front \leq i \leq n $.
 
During the loop: There are three possible scenarios at each iteration:
 
Case 1: Tail is selected such that $\alpha < tail$. In this case $back$ is set to $ind$. Assume there exists $\tau\opt $ such that $\tau\opt >= $x[ind]$ $, and $x[\tau\opt]=\quant^+$. Then $P[x < x[\tau\opt]] \geq tail > \alpha$, meaning that $x[\tau\opt]$ is not feasible by \eqref{eq:var-definition}. %However, if $\tau\opt $ such that $\tau\opt \leq ind$ , then $P[x < x[\tau\opt]]  < tail > \alpha$, meaning that $x[\tau\opt]$ is potentially feasible by \eqref{eq:var-definition}, so by Lemma 1 $\quant^+$ must remain in this range.% 
 Since from Lemma 1 we know $\quant^+$ must be in $x$, the loop invariant must hold true.

Case 2: Tail is selected such that $\alpha > tail$. In this case $front$ is set to $ind + 1$. Assume there exists $\tau\opt $ such that $\tau\opt \leq $x[ind]$ $, and $x[\tau\opt]=\quant^+$. Let $\tau' = ind$. $P[x < x[\tau']] \leq tail < \alpha$. If $x[\tau\opt] <  x[\tau']$ then $x[\tau\opt]$ can not be optimal by \eqref{eq:var-definition}. If $x[\tau\opt] =  x[\tau']$, then from the partitioning scheme we this is the optimal $\quant^+$ since 
 
 Case 3: $front - back = 0$.
 Two scenarios here. If $x$ contains more than one element, all the elements are no longer in the range $[front, back]$ could not have been $\quant_{\alpha}^+(\tilde{x})$ (see Case 1 and Case 2), and $\quant_{\alpha}^+(\tilde{x})$ must exist in x per lemma 1. Therefore this element that will be returned is $\quant_{\alpha}^+(\tilde{x})$.  If $x$ had one element to begin with, $\quant_{\alpha}^+(\tilde{x})$ equals this element for all admissible alpha. 

 Termination Step:  
 Proof: At every iteration, the algorithm will always hit case 1 or case 2 assuming $back - front >= 1$. If $alpha < tail$, $back$ will always get smaller assuming it is in the range $(front, back$. If $alpha \geq tail$ , then  $back - front$ will not get smaller if $ind = front$. However, in this case the elements at $front$ and $back$ will have been swapped in the partition scheme, such that at next iteration $ind$ will not be equal to $front$ (as the pivot selected will now no longer be the smallest element in the list). 
\end{proof}

\begin{lem}
\Cref{alg:quick_quantile}in \eqref{eq:var-definition} runs in linear time.
\end{lem}
\begin{proof}
  \mm{Maeve and prove it in Lean}
\end{proof}
\begin{lem}
CVAR can be computing by solving an optimization problem $\max_{Q \in \mathcal{Q}} \mathbb{E}_Q[X]$, such that all $Q$ in the set must satisfy such that all  $Q \in \mathcal{Q}  satisfy \max_{Q} Q(A) < \frac{1}{1-\alpha} P(A) \quad \forall A \in \mathcal{P}(\Omega)$ . 
\end{lem}
\begin{proof}
  We start by noting that Q(A) can be written as $\sum_{j=1}^{n} \xi_j(A_j)$ where $A_1 , A_2, .. ,A_n$, are disjoint sets corresponding to a outcome where only one possible value of the random variable is realized and A is a subset of $\Omega$ if $j \neq n$. For all Q(A) such that $j \neq n$, we know from (3) that Q(A) \leq$\sum_{j=1}^{n} (1/1-a)*P(a)$ and that that $Q(A) < 1$ , thus Q(A) \leq $min($\sum_{j=1}^{n} (1/1-a)*P(a), 1)$. 

\end{proof}
\begin{lem}
We can prove that for any given $A$ we can construct a feasible $Q$ such that $Q(A) = min($\sum_{j=1}^{n} (1/1-a)*P(a), 1)$.

\end{lem}
\begin{proof}
  Two possible scenarios exist. 
  Case 1: $P(A) \geq 1 - \alpha$. In this case, Q(A) must equal 1 since $(1/(1 - \alpha))* P(A) > 1 $. Thus every element in $A$ should be scaled by $1/P(A)$ if in $A$ and set to 0 otherwise. Since  $P(A) \geq 1 - \alpha$, $1/P(A) \leq 1/(1-\alpha)$  
  Case 2: $P(A) < (1 - \alpha)$. In this scenario, $Q(A) \leq (1/(1 - \alpha)) * P(A) < 1$. For any given A, We can construct a Q(A) such that $Q(A) = (1/(1 - \alpha)) * P(A)$. This can be done by setting $Q(a) to $(1/(1 - \alpha))* P(a)$ for all a in A and $P(a)* d$ otherwise such that $(1/(1 - \alpha)) * P(A) + d*P(B) = 1$, where B contains all elements not in A. It is provable that $d \leq (1/(1 - \alpha))$, which is required for the constructed Q to be feasible. For all feasible $\alpha \geq 0$, $(1/(1 - \alpha)) \geq 1. Since $P(A) + P(B) = 1$,  $d must be \leq 1$ in order for $(1/(1 - \alpha)) * P(A) + d*P(B) = 1$ to hold. Since $1 \leq 1/(1-\alpha)$, d must then be \leq (1/(1 - \alpha)).
\end{proof}

\begin{lem}
  If \cref{alg:quick_quantile} correctly computes $\quant^+$ then \cref{alg:quick_cvar} correctly computes CVaR. 
\end{lem}
\begin{proof}
  \mm{TODO}
\end{proof}

\section{Algorithm}

\subsection{Piece-wise Linear Minimization}
\maeve{Incorporate projection.jl }

\subsection{Quick Quantile}

\begin{algorithm}
  \algrenewcommand\algorithmicrequire{\textbf{Input: Random variable $\tilde{X}$ as vector $x$ and probability mass function $p$ and $\alpha \in [0,1]$}}
  \algrenewcommand\algorithmicensure{\textbf{Output: $\alpha$-quantile of $\tilde{X}$}}
  \caption{algorithm}{Quick Quantile}\label{alg:quick_quantile}
  \begin{algorithmic}[1]
    \Require{}
    \Ensure{}
    \todo{Make absolutely sure that you tell the reader that we use 1based indexing}
    \State{$front \gets 1$}
    \State{ $back \gets n$ }
    \While{$back - front \geq 1$}
    \State{$ind \gets partition(x, p, f, b) - 1$  \Comment{Lomuto partition scheme - works in the code but termination can not be proven}}
    \State{$tail \gets \sum_{i=1}^n \mathds{1}(x[i] \leq x[ind]) p[i]$}
      \If{$\alpha < tail$}
      \State{$back \gets ind$}
      \Else{}
      \State{$front \gets ind + 1$}
      \EndIf{}
    \EndWhile{}
    \State{\Return{$x[front]$}}
  \end{algorithmic}
\end{algorithm}


\subsection{Quick CVaR}

\begin{algorithm}
  \floatname{algorithm}{Quick CVaR}
  \algrenewcommand\algorithmicrequire{\textbf{Input: Random variable $\tilde{X}$ as vector $x$ with probability mass function p and $\alpha \in [0,1]$}}
  \algrenewcommand\algorithmicensure{\textbf{Output: $CVaR_\alpha(\tilde{X})$}}
  \caption{}\label{alg:quick_cvar}
  \begin{algorithmic}[1]
    \Require{}
    \Ensure{} 
    \todo{Make absolutely sure that you tell the reader that we use 1 or 0 based indexing}
    \State{$q^+_\alpha \gets \text{quick\_quantile}(x, p, \alpha)$}
    \State{$val \gets 0$}
    \State{$p_{left} \gets 1$}
    \For{$i = 0 \dots n$}
      \State{$inc \gets \min(\frac{p[i]}{\alpha}, p_{left})$}
      \State{$val += x[i] * inc$}
      \State{$p_{left} -= inc$}
      \If{$p_{left} \leq 0$}
        \State{Break}
      \EndIf{}
    \EndFor{}
    \State{\Return{$val$}}
  \end{algorithmic}
\end{algorithm}



\begin{algorithm}
\floatname{algorithm}{Partition}
\algrenewcommand\algorithmicrequire{\textbf{Input: Random variable $\tilde{X}$ as vector $x$ with probability mass function p and $lo, hi \in \mathbb{R}$}}
\algrenewcommand\algorithmicensure{\textbf{Output: pivot index $i \in \mathbb{R}$}}
  \begin{algorithmic}[1]
    \Require{}
    \Ensure{} 
    \State{pind = f + (b-f)/2}
    \State{pivot = x[pind]}
    \State{swap x[pind] with x[hi]}
    \State{swap p[pind] with p[hi]}
    \State{i = lo}
    \For{$j = lo \dots hi - 1$}
        \If{x[j] $\leq$ pivot}
          \State{swap x[i] with x[j]}
          \State{swap p[i] with p[j]}
          \State{i += 1}
        \EndIf{}
    \EndFor{}
    \State{swap x[i] with x[hi]}
    \State{swap p[i] with p[hi]}
    \State{\Return{i \Comment{the pivot index}}}
  \end{algorithmic}
\end{algorithm}


\section{k-capped simplex}

\begin{equation}
  \label{eq:k_capped_simplex}
  \psi(x, k) = 
  \begin{mprog}
    \minimize{\xi \in \Real^n} \lvert \lvert \xi - x \rvert \rvert_2^2
    \stc \xi \in \Delta_k^n
  \end{mprog}
  =
  \begin{mprog}
    \minimize{\bm\xi \in \Real^n} \lvert \lvert \bm\xi - \bm{x} \rvert \rvert_2^2
    \stc \bm\xi\tr \mathbf{1} \leq k
    \cs \bm{0} \le \bm\xi \le \bm{1} 
  \end{mprog}
\end{equation}

\mm{define k-capped simplex}

    \todo{Make absolutely sure that you tell the reader that we use 1based indexing. Make sure that we always use 1-based indices in all algorithms}

\begin{algorithm}
  \algrenewcommand\algorithmicrequire{\textbf{Input:   $\tilde{X}$ as vector $x$ to project and sum constraint k}}
  \algrenewcommand\algorithmicensure{\textbf{Output: $\alpha$-quantile of $\tilde{X}$}}
  \caption{algorithm}{K-capped}\label{alg:quick_quantile}
  \begin{algorithmic}[1]
    \Require{}
    \Ensure{}
    \State{$frontA \gets 1$}
    \State{ $backA \gets n$ }
    \State{$frontB \gets 1$}
    \State{ $backB \gets n$ }
    \State{ $midSum \gets 0$ }
    \State{ $minMid \gets inf$ }
    \State{ $midSize \gets 0$ }
    \State{ $maxMid \gets -inf$ }
    \While{$backB - frontB \geq 1$ and $backA - frontA \geq 1$}
      
     \Comment{Lomuto partition scheme}
    
      \If{$backB - frontB >= 1$}
      \State{$ind,leftsum,maxLeft,minLeft \gets modifiedPartition(x, l, f, b) - 1$}
      \State{$\gamma \gets -2x[ind] + 2$}
       \If{$maxMid + \gamma*2 >= 1$}
         \State{$front \gets ind + 1$}
       \ElsIf{$minMid + \gamma*2 <= 0$}
         \State{$back \gets ind$} 
       \Else{}
       \State{$sum = \sum_{n=frontA}^{backA} (max(0, min(1, x[i])))  + midSum + \gamma*midSize + \sum_{i=frontB}^{backB} (max(0, min(1, x[i])))$}
       \If{$sum > k$}
           \State{$back \gets ind - 1$}
       \ElsIf {$sum < k$}
           \State{$front \gets ind$}
           \If {$frontB - backA > midsize$}
           \State{$midSize \gets frontB - backA$}
           \State{$midSum \gets leftSum + midSum$}
            \EndIf{}
            
           
       \Else{}
         return solution with computed $\gamma$
        
      \EndIf{}
      
      \EndIf{}
      \EndIf{}
    \If{$backA - backB \ge 1$} 
       \State{ $backA \gets backB$ }
     \EndIf{}
    \If{$backA - frontA >= 1$}
      \State{$ind,rightsum,maxRight,minRight \gets modifiedPartition(x, r, f, b) - 1$}
      \State{$\gamma \gets -2x[ind]$}
       \If{$maxMid + \gamma*2 >= 1$}
         \State{$front \gets ind + 1$}
       \ElsIf{$minMid + \gamma*2 <= 0$}
         \State{$back \gets ind$} 
       \Else{}
       \State{$sum = \sum_{n=frontA}^{backA} (max(0, min(1, x[i])))  + midSum + \gamma*midSize + \sum_{i=frontB}^{backB} (max(0, min(1, x[i])))$}
       \If{$sum > k$}
           \State{$back \gets ind - 1$}
       \ElsIf {$sum < k$}
           \State{$front \gets ind$}
           \If {$frontB - backA > midSize$}
           \State{$midSize \gets frontB - backA$}
           \State{$midSum \gets rightSum + midSum$}
            \EndIf{}
            
           
       \Else{}
         
          \EndIf{}   
     \EndIf{}
      \EndIf{}  
  \EndWhile{}
    \State{\Return{$x[front]$}}

  \end{algorithmic}
\end{algorithm}

\section{Greedy Algorithm}

\begin{algorithm}
    \algrenewcommand\algorithmicrequire{\textbf{Input: }Finite $N = [1 \ldots n]$, non-decreasing submodular function $f : 2^N \to \mathbb{R}$ satisfying $f(\emptyset) = 0$ and $f(N) = 1$} 
    \algrenewcommand\algorithmicensure{\textbf{Output: $\max \{f(S) : S\subseteq N\}$}}
  \caption{algorithm}{Greedy algorithm from Submodular Functions, Polymatroids, TDI}
 \begin{algorithmic}[1]
    \Require{}
    \Ensure{}
    \State{Order the elements such that $c_1 \geq c_2 \cdots \geq c_n$.}
    \State{Let $S_i := \{1, \dots, i\}$ and $S_0 := \emptyset$}
    \State{Let $x_i :=  \begin{cases}f(S_i) - f(S_{i-1}) & \text{for } i = 1,\dots, r \\ 0 & \text{otherwise} \end{cases}$}
\end{algorithmic}
\end{algorithm}

\section{Comonotonicity of Risk}

Given two random variables $\tilde{x}$ and $\tilde{y}$ we say that they are comonotonic if \todo{what is the definition of comonotonicity?}.
A risk measure is said to be comonotonic if $\rho(\tilde{x} + \tilde{y}) = \rho(\tilde{x}) + \rho(\tilde{y})$ for comonotonic random variables $\tilde{x}$ and $\tilde{y}$.

\begin{table}[]
\begin{tabular}{lll}
  \textbf{Risk Measure} &  \textbf{Comonotone?} & \\
  $\cvaro$ & Yes & \\
  $\varo$ &  Yes & \\
  Un weighted L1 &  Yes & \\
  $\evaro$ & No & \\
  Expectile & No &
\end{tabular}
\end{table}

\bibliography{fastcvar.bib}

\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End:
